---
title: "Synthetic data and LMTP demo"
author: 'Aimee Harisson and Andy Wilson'
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(tmle)
library(ltmle)
data(sampleDataForLtmleMSM)
library(lmtp)

```


```{r}
# install.packages("ltmle_1.2-0.tar.gz", repos = NULL, type = "source")
```

# Thanks

Special Thanks to Haodong Li! 

# Libraries used

```{r}
library(tidyverse)
library(tmle)
library(ltmle)
data(sampleDataForLtmleMSM)
library(lmtp)
```


# Generate synthetic data


```{r}
# data with two time stage: O = (L(1), A(1), L(2), A(2), Y)
get_data  <- function(n) {
  L1 = rnorm(n)
  A1 = rbinom(n, 1, plogis(0.01*(L1)))
  L2 = 0.5*L1 + 0.7*A1 + rnorm(n)
  A2 = rbinom(n, 1, plogis(0.01*(L2+A1+L1)))
  Y = rbinom(n, 1, plogis(3*L1*A1 + 3*L2*A2))
  O = data.frame(L1, A1, L2, A2, Y)
  return(O)
}
```



##  helper function generating counterfactual data

```{r}
get_data_d <- function(n, abar = 1) {
  L1 = rnorm(n)
  if (abar == "DTR") {
    A1 = as.numeric(L1 < 0)
  } else {
    A1 = abar[1]
  }
  L2 = 0.5*L1 + 0.7*A1 + rnorm(n)
  if (abar == "DTR") {
    A2 = as.numeric(L2 < 0)
  } else {
    A2 = abar[1]
  }
  Y = rbinom(n, 1, plogis(3*L1*A1 + 3*L2*A2))
  O = data.frame(L1, A1, L2, A2, Y)
  return(O)
}
```



## verify truth 

$E(Y_{\bar{a} = 1})$ = 0.64

```{r}
set.seed(123)
df_big <- get_data_d(1e6, abar = 1)
mean(df_big$Y)
```


## verify truth (policy) 

$E(Y_{\bar{a} = d})$ = 0.26, where d = $L1 < 0, L2 <0$

```{r}
df_big <- get_data_d(1e6, abar = "DTR")
mean(df_big$Y)
```



# generate observed data

```{r}
set.seed(123)
df <- get_data(1e4)
head(df)
```


# ltmle estimation(s)



## ltmle estimate 1: always on

```{r}
res.11 = ltmle(df,
            Anodes = c("A1", "A2"),
            Lnodes = c("L1", "L2"),
            Ynodes = "Y",
            abar = c(1,1),
            SL.library = c("SL.glm", "SL.earth", "SL.xgboost"), 
            SL.cvControl = list(V=10))
summary(res.11)

```



## ltmle estimate 2:  always off

```{r}
res.00 = ltmle(df,
            Anodes = c("A1", "A2"),
            Lnodes = c("L1", "L2"),
            Ynodes = "Y",
            abar = c(0,0),
            SL.library = c("SL.glm", "SL.earth", "SL.xgboost"), 
            SL.cvControl = list(V=10))
summary(res.00)
```




## ltmle estimate 3: dynamic rule

```{r}
d <- cbind(df$L1 < 0, df$L2 <0)

res_d = ltmle(df,
              Anodes = c("A1", "A2"),
              Lnodes = c("L1", "L2"),
              Ynodes = "Y",
              abar = d,
              SL.library = c("SL.glm","SL.earth", "SL.xgboost"), 
              SL.cvControl = list(V=10))
summary(res_d)
```



## ltmle estimate 4:  ATE

```{r}
res.ATE = ltmle(df,
            Anodes = c("A1", "A2"),
            Lnodes = c("L1", "L2"),
            Ynodes = "Y",
            abar = list(c(1,1), c(0,0)),
            SL.library = c("SL.glm", "SL.earth", "SL.xgboost"), 
            SL.cvControl = list(V=10))
(Z <- summary(res.ATE))
```

### Odds ratio comparing always on vs. always off

```{r}
Z$effect.measures$OR$estimate
Z$effect.measures$OR$CI
```


# Bonus: library(lmtp)



```{r}
# our treatment policy function to be applied at all time points
policy <- function(data, trt) {
  if (trt == "A1") {
    return(as.numeric(data$L1 < 0))
  } else if (trt == "A2") {
    return(as.numeric(data$L2 < 0))
  }
}
```


In addition to specifying a treatment policy, we need to specify our treatment variables and time-varying covariates.

```{r}
# treatment nodes
A <- c("A1", "A2")
# time varying nodes
L <- list(c("L1"), c("L2"))
```

We can now estimate the effect of our treatment policy, `d`. 

In this example, we’ll use the cross-validated TML estimator with 10 folds.



```{r}
result <- lmtp_tmle(
  df, 
  A, 
  "Y", 
  time_vary = L, 
  shift = policy, 
    # intervention_type = "mtp", 
    # intervention_type = "det", new format: mtp = FALSE
  mtp = FALSE,
  folds = 10
)

print(result)

# Unfortunately, this currently prints as a 'warning' 
# so needs to be manually pasted: 

# LMTP Estimator: TMLE
#    Trt. Policy: (policy)
# 
# Population intervention estimate
#       Estimate: 0.268
#     Std. error: 0.0028
#         95% CI: (0.2625, 0.2735)
```

```{r}
result$theta
result$standard_error
c(result$low, result$high)
```

